---
title: "Access and visualize FDA food recall data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Access and visualize FDA food recall data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

foodRecall is an R package that facilitates accessing and visualizing data from the openFDA API. This document provides descriptions and examples of the functions included in the package and some information explaining the process to complete this project for STAT 585 - Data Technologies for Statistical Analysis at Iowa State University.

## Background/Motivation

## Functions and applications

```{r setup}

library(foodRecall)

# Before running any of the package functions, the user will need an API key.
# To get the API key, go to https://open.fda.gov/apis/authentication/. It is free, only requires
# an email address, and you will get it immediately.


# Get food recall locations
# recall_location() - this function uses search criteria such as api_key, city, and state.

# df <- foodRecall::recall_location(api_key = api_key, limit = 1000, city = "Kansas City, Ames", state = "Iowa, Missouri")


# Map food recall locations
# map_recall()

# map <- foodRecall::map_recall(data = df)

# Working with food recall dates
# recall_date()


```

## Limitations 

## Future Work


